\clearpage
\pagebreak

\section*{Abstract}
This thesis, entitled \textit{"Procedural terrain generation for underwater environments"}, as its name suggests, focuses on the topic of procedural terrain generation with the specificity to tackle the tackle underwater environments. Terrain generation is still an open research area in computer graphics as, with the emergence of simulation, rendering, and interaction techniques, the entire field is experiencing increased collaboration with terrain experts. 
Virtualization of the physical world helps users to see and manipulate it, allowing for a better understanding of it and bringing together numerous experts, gradually breaking down the boundaries of scientific disciplines. 
% Bringing the physical world into a computer and allowing users to see and manipulate this virtual world helps to better understand it and to bring together numerous experts, gradually breaking down the boundaries of scientific disciplines. 
Terrain science brings together, in almost a direct manner, geologists, oceanologists, physicists, meteorologists, biologists, roboticists, computer scientists, 3D artists, and more. We focused our work on the inclusion of the user in the generation process of virtual worlds through fast and controlable algorithms, with the possibility to dive underwater.

This thesis is divided into three parts, guiding the user from the design of a landscape, its 3D modeling, until its finalization, step by step.
Firstly, we will propose a formalization of terrain designing, allowing for the conception of environments in a semantic sense, abstracting away from 3D geometry and data structures.

Secondly, we will see how to give 3D form to these environments. We will propose new models for the generation and modeling of coral islands and karst networks.

In the third part, we will focus on adding realism to 3D terrains through physical simulations of erosion processes. We will demonstrate a flexible and controllable method to imitate the long-term effects of water and wind on both terrestrial and marine landscapes.

\textbf{Keywords:} terrain representation, procedural generation, physical simulations, user interaction

% \section*{Résumé}
% Cette thèse porte sur le thème de la génération procédurale de terrains en milieux sous-marins. La génération de terrain est un domaine de recherche encore ouvert en informatique graphique car, avec l'emergence des techniques de simulation, de rendu et d'interaction, le domaine entier connait un accroissement de collaboration avec les experts terrains. Amener le monde physique dans un ordinateur et permettre à son utilisateur de voir et manipuler ce monde virtuel permet de mieux le comprendre et de réunir de nombreuses experts ensemble, brisant petit à petit les frontières des disciplines scientifiques. La science des terrains rassemble, de manière presque directe, géologues, oceanologues, physiciens, météorologues, biologistes, roboticiens, informaticiens, artistes, etc... 

% Cette thèse se divise en trois parties amenant l'utilisateur de la conception d'un paysage jusqu'à sa finalisation, étape par étape, en gardant le control en tout point. 
% Premièrement, nous proposerons une formalisation d'esquissage de terrains, permettant la conception d'environnements dans un sens sémantique, permettant de s'abstraire de la géometrie et de structures de données. 

% Secondement, nous verrons comment donner forme 3D à ces esquisses. Nous proposerons de nouveaux algorithmes pour la génération et modélisation d'iles coralliennes et de réseaux karstiques.

% En troisième partie, nous nous interesserons à l'ajout de réalisme sur terrains existants au travers de simulations physiques de processus d'érosion. Nous montrerons une méthode flexible et controlable pour imiter les effets de l'eau et du vents à très long terme sur un paysage terrestre et marin.

% \textbf{Mots-clés :} représentation de terrain, génération procédurale, simulations physiques, interaction utilisateur