\part{Modelisation}

\chapter*{Abstract}
\label{chap:modelisation-abstract}

Once the global shape of the environment is defined by the designer, the modeling of the surface may begin. Modeling a terrain allows us to visualize the surface of the terrain and see concretely the result of the generation process. In the modeling process of landscape features, the method used for each of the element may be completely different as the natural phenomena forming or creating them may have very different natures. With procedural generation, the balance between speed, realism and control have to be stroke. For this reason, the same geological feature may be simulated using completely different algorithms.

The modeling part of this thesis is divided in two chapters, each of them focusing on a different landscape feature. In \cref{chap:coral-island}, we will propose a novel algorithm for the modeling of coral islands taking advantage of the recent advances in deep learning. We will present our method for generating a large dataset of examplar and the training of a Generative Adversarial Network (GAN).

Secondly in \cref{chap:karsts}, we describe a procedural algorithm for the modeling of karst networks. In this method, we define the path of the cavities, focused toward user control.



% - Methods completely different \\
% ** Physical phenomena are different \\
% ** => Simulation/generation methods must be specific for one landscape \\
% - Methods are developed at different instants of the thesis work, \\
% ** We will see different procedural generation domains: \\
% *** Analytical solutions (coral islands \#1) \\
% *** Deep Neural Networks (coral islands \#2) \\
% *** 3D user interaction (karst networks) \\
% - Deep Learning tends to replace procedural methods in 2D domain, \\
% ** But still too complex for 3D models \\
% *** (lack of data, lack of research interest for now) \\
% ** Still requires a lot of data, which, is (while not easy) possible using aerial images, but is way too sparse with underwater landscapes or underground biomes. \\
% - We want to control the area in which an element is modeled in the terrain \\
% ** Because that's how we define them in the previous chapter. \\
% ** Thus we cannot use simple random noise methods => no bounds \\
% *** Only solution is to use falloff maps, but meh... \\
% - As such, we want to keep the skeleton of our elements using primitives (points, curves, regions) \\
% ** => Much easier to add constraints / manipulate primitives in a "procedural generation" way. \\
% - Warning: usage of Deep Learning is at the limit of "procedural generation" and is not considered as part of it by the whole community. \\
% (Complete with the Reddit poll). \\
% - ... 


% \chapter{Graphical representation of environmental objects}
% \label{chap:representation-env-objects}
% - Implicit surfaces \\
% - Meshes \\
% - ...


% \input{Chapter 2/ModelisationVolumique}

\input{Chapter 2/GenerationIlesCoraliennes}

\input{Chapter 2/GenerationKarsts}